

## quality control metrics

```{r load-metrics}
qc = read.table(file.path(path_results, "metrics", "metrics.tsv"),
                header=T, sep="\t", check.names=F,
                colClasses=list("sample"="character"))
rownames(qc) = qc$sample
qc$Mapped_reads_pct = as.numeric(gsub("%", "", qc$Mapped_reads_pct))
qc$Duplicates_pct = as.numeric(gsub("%", "", qc$Duplicates_pct))
```

### Mean quality along read

```{r read-qual}
qual = read.table(file.path(path_results, "fastqc", "Per_base_sequence_quality.tsv"), header=T, sep= "\t", check.names =T, colClasses=list("sample"="character"))
qual$sample = as.character(qual$sample)

ggplot(qual, aes(x=Base, y=Median, group=sample)) +
    geom_line() +
    geom_line(aes(x=Base, y=Lower_Quartile, group=sample, col="lower_quantile")) +
    geom_line(aes(x=Base, y=Upper_Quartile, group=sample, col="upper_quantile")) +
    geom_line(aes(x=Base, y=X10th_Percentile, group=sample, col="10%_quantile")) +
    facet_wrap(~sample) +
    ylim(0,41) +
    scale_color_brewer("metrics",palette = "Set1") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5))+
    scale_x_discrete(breaks=number_ticks(10))
```

### Size distribution

```{r read-size}
qual = read.table(file.path(path_results, "fastqc", "Sequence_Length_Distribution.tsv"), header=T, sep= "\t", check.names = F, colClasses=list("sample"="character"))
qual = qual %>% group_by(sample) %>% mutate(total=sum(Count), pct=Count/total)

ggplot(qual , aes(x=Length, y=Count, group=sample)) +
    geom_line(size=2, alpha=0.5) +
    scale_color_brewer(palette = "Set1") +
    theme_bw() +
    facet_wrap(~sample) +
    labs(y="# of reads")+
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5)) +
    scale_x_discrete(breaks=number_ticks(5))
```

Length read with more than 25% of the total reads.

```{r table-size, results='asis'}
kable(qual %>% filter(pct > 0.25) %>% dplyr::select(Length, sample, pct) %>% spread(Length, pct), align="c", digits=2)

```

### GC content

Samples where less than 80% of the reads are inside 30%-70% GC content.

```{r table-gc, results='asis'}
qual = read.table(file.path(path_results, "fastqc", "Per_sequence_GC_content.tsv"), header=T, sep= "\t", check.names = F, colClasses=list("sample"="character"))
qual = qual %>% mutate(pct=Count/total) %>% filter(GC_Content>10 & GC_Content<90) %>% 
   group_by(sample) %>% summarise(pct_in_30_70=sum(pct))

kable(qual %>% filter(pct_in_30_70 < 0.95))

```


### Nucleotide content

```{r read-content}
qual = read.table(file.path(path_results, "fastqc", "Per_base_sequence_content.tsv"), header=T, sep= "\t", check.names = F, colClasses=list("sample"="character"))
qual$sample = as.character(qual$sample)

qual$Base = as.numeric(qual$Base)

dd = melt(qual, id.vars = c("Base", "sample"), variable_name = c("nt"))

ggplot(dd, aes(x=Base, y=value, group=sample)) +
        geom_line(size=2, alpha=0.5) +
        theme_bw() +
        ylab("% of nucleotides") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5)) +
        ylim(10,50) +
    facet_wrap(~nt)


```

Nt in position with > 30% or < 10% of total reads.

```{r table-content, results='asis'}
kable( melt(qual, id.vars=c("Base", "sample"), variable_name = "nt") %>%
           filter(value > 30 | value < 10) %>% filter(Base<20) %>%
           dplyr::select(Base, sample, nt, value) %>%
           spread(Base, value),
       align="c", digits=2)
```


### Total reads

```{r total-reads}
ggplot(qc, aes(x=sample, y=Total_reads/1e6)) +
    geom_bar(stat = 'identity') +
    ylab("Million reads") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Mapped reads

```{r mapped-reads}
ggplot(qc, aes(x=sample, y=Mapped_reads/Total_reads)) +
    geom_bar(stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Offtarget reads

```{r off-reads}
ggplot(qc, aes(x=sample, y=offtarget/Mapped_reads)) +
    geom_bar(stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


### Basic metrics by samples

```{r table, results='asis'}
metrics = c("sample", "Total_reads" ,"Mapped_reads_pct", "Duplicates_pct",
            "offtarget",
            "%GC", "Sequence_length", "Median_insert_size")
qc$offtarget = qc$offtarget/qc$Total_reads
# qc$secondary = qc$secondary/qc$Total_reads
# qc$singletons = qc$singletons/qc$Total_reads
print(kable(qc[, metrics], align="c", digits=2))

```

### Variant metrics by samples

```{r table-variants, results='asis'}
qc$ratio_het_hom = qc$Variations_heterozygous/qc$Variations_homozygous
metrics = c("sample", "Variations_total", "Variations_in_dbSNP_pct",
            "Variations_heterozygous", "Variations_homozygous", "ratio_het_hom", "Transition/Transversion")
print(kable(qc[, metrics], align="c", digits=2))

```

