
# Basic sample metrics

A quick summary of sample metrics to identify outliers. Offtargets will be set to 0 for non-targeted experiments, and table cells are color-coded to indivate deviation from the mean for a given metric. Note that this may be expected depending on the experimental setup.

```{r table, results='asis'}
qc = read.table(file.path(path_results, "metrics", "metrics.tsv"),
                header=T, sep="\t", check.names=F,
                colClasses=list("sample"="character"))
rownames(qc) = qc$sample

# Define metrics to display and fix missing
metrics = c("sample", "Total_reads" ,"Mapped_reads_pct", "Duplicates_pct",
            "offtarget", "%GC", "Sequence_length", "Average_insert_size")
missing = setdiff(metrics, colnames(qc))
if ( length(missing) >0 ){
    for (col in missing){
        qc[, col] = 0
    }
}
if ("Median_insert_side" %in% colnames(qc))
    qc[,"Average_insert_size"] = qc[,"Median_insert_size"]

if (is.null(qc$offtarget)) {
  qc$offtarget <- rep  (0, length(qc$sample))
} else {
  qc$offtarget = qc$offtarget/qc$Total_reads
}

# Adjust some of the text information for formatting purposes
qc_table <- qc[, metrics]

qc_table$Mapped_reads_pct = as.numeric(gsub("%", "", qc_table$Mapped_reads_pct))
qc_table$Mapped_reads_pct <- qc_table$Mapped_reads_pct / 100

qc_table$Duplicates_pct = as.numeric(gsub("%", "", qc_table$Duplicates_pct))
qc_table$Duplicates_pct <- qc_table$Duplicates_pct / 100

# Init intervals for coloring
total_reads_intrv = quantile(c(0,qc_table$Total_reads), c(.25,.50,.75))
mapped_reads_intrv = quantile(c(0,qc_table$Mapped_reads_pct), c(.25,.50,.75))
duplicated_intrv = quantile(c(0,qc_table$Duplicates_pct), c(.25,.50,.75))
gc_intrv = c(.1,.30,.70, .9) * 100
offtarget_intrv = c(.10,.30,.60)
duplicated_intrv = c(.10,.30,.60)
fake_insert_size = rnorm(10,mean(qc_table$Average_insert_size))
insert_size_intrv = c(mean(fake_insert_size) - 2 * sd(fake_insert_size),
                      mean(fake_insert_size) - sd(fake_insert_size),
                      mean(fake_insert_size) + sd(fake_insert_size),
                      mean(fake_insert_size) + 2 * sd(fake_insert_size))

if (nrow(qc_table) > 7){
    total_reads_intrv = c(mean(qc_table$Total_reads) - 2 * sd(qc_table$Total_reads),
                          mean(qc_table$Total_reads) - sd(qc_table$Total_reads),
                          mean(qc_table$Total_reads))
    mapped_reads_intrv = c(mean(qc_table$Mapped_reads_pct) - 2 * sd(qc_table$Mapped_reads_pct),
                           mean(qc_table$Mapped_reads_pct) - sd(qc_table$Mapped_reads_pct),
                           mean(qc_table$Mapped_reads_pct))
    duplicated_intrv = c(mean(qc_table$Duplicates_pct),
                         mean(qc_table$Duplicates_pct) + sd(qc_table$Duplicates_pct),
                         mean(qc_table$Duplicates_pct) + 2 * sd(qc_table$Duplicates_pct))
    gc_intrv = c(mean(qc_table$"%GC") - 2 * sd(qc_table$"%GC"),
                 mean(qc_table$"%GC") - sd(qc_table$"%GC"),
                 mean(qc_table$"%GC") + sd(qc_table$"%GC"),
                 mean(qc_table$"%GC") + 2 * sd(qc_table$"%GC"))
    offtarget_intrv = c(mean(qc_table$offtarget),
                        mean(qc_table$offtarget) +  sd(qc_table$offtarget),
                        mean(qc_table$offtarget) + 2 * sd(qc_table$offtarget))
    insert_size_intrv = c(mean(qc_table$Average_insert_size) - 2 * sd(qc_table$Average_insert_size),
                          mean(qc_table$Average_insert_size) - sd(qc_table$Average_insert_size),
                          mean(qc_table$Average_insert_size) +i sd(qc_table$Average_insert_size),
                          mean(qc_table$Average_insert_size) + 2 * sd(qc_table$Average_insert_size))
}

# Calculate mean and SD where appropriate
datatable(qc_table,
          rownames=FALSE,
          options=list(dom = 'tp',
                       pageLength = length(qc_table$sample))) %>%
  formatPercentage('Mapped_reads_pct', 1) %>%
  formatPercentage('Duplicates_pct', 1) %>%
  formatPercentage('offtarget', 1) %>%
  formatStyle('Total_reads',
              backgroundColor = styleInterval(total_reads_intrv,
                                              c('#f03b20',
                                                '#feb24c',
                                                '#ffeda0',
                                                'white'))) %>%
  formatStyle('Mapped_reads_pct',
              backgroundColor = styleInterval(mapped_reads_intrv,
                                              c('#f03b20',
                                                '#feb24c',
                                                '#ffeda0',
                                                'white'))) %>%
  formatStyle('Duplicates_pct',
              backgroundColor = styleInterval(duplicated_intrv,
                                              c('white',
                                                '#ffeda0',
                                                '#feb24c',
                                                '#f03b20'
                                                ))) %>%
    formatStyle('%GC',
              backgroundColor = styleInterval(gc_intrv,
                                              c('#feb24c',
                                                '#ffeda0',
                                                'white',
                                                '#ffeda0',
                                                '#feb24c'
                                                ))) %>%
  formatStyle('offtarget',
              backgroundColor = styleInterval(offtarget_intrv,
                                              c('white',
                                                '#ffeda0',
                                                '#feb24c',
                                                '#f03b20'
                                                ))) %>%
  formatStyle('Average_insert_size',
              backgroundColor = styleInterval(insert_size_intrv,
                                              c('#feb24c',
                                                '#ffeda0',
                                                'white',
                                                '#ffeda0',
                                                '#feb24c'
                                                )))
```

## Total and mapped reads

The next two plots compare the number of reads in each sample (should be uniform) and the percentage of reads mapping to the reference genome. Low mapping rates are indicative of sample contamination, poor sequencing quality or other artifacts.

```{r total-reads}
ggplot(qc, aes(x=sample, y=Total_reads/1e6)) +
    geom_bar(stat = 'identity') +
    ylab("Million reads") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r mapped-reads}
ggplot(qc, aes(x=sample, y=Mapped_reads/Total_reads)) +
    geom_bar(stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

The _effective_ number of reads is the total number of mapped read adjusted for duplicates; a high duplicate rate can reduce the overall coverage that will contribute to variant calling processes:

```{r effective-reads}
ggplot(qc, aes(x=sample, y=(Total_reads-Duplicates)/1e6)) +
    geom_bar(stat = 'identity') +
    ylab("Million reads") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Offtarget reads

In addition to total read count and map-ability it is important to know if the target enrichment (if any) worked, i.e., what percentage of reads are on the amplified or captured regions, and what percentage is considered 'off target'. On target percentages should be uniform and ideally above 50% for most capture methods. 
```{r off-reads}
ggplot(qc, aes(x=sample, y=offtarget*100)) +
    geom_bar(stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Sequencing metrics

Various metrics related to the quality of the sequencing data itself, i.e., the FASTQ or BAM fie as it came off the sequencer. 

## Mean sequence quality along read

Mean base quality for each sample along with the upper and lower quartile, and the 10% (lower) quantile. A slight dropoff towards the end of the read is normal, but you want the mean qualities and quartiles to track each other closely with relatively few outliers. In addition, all plots should look similar to each other.


```{r read-qual}
qual = read.table(file.path(path_results, "fastqc", "Per_base_sequence_quality.tsv"), header=T, sep= "\t", check.names =T, colClasses=list("sample"="character"))
qual$sample = as.character(qual$sample)

ggplot(qual, aes(x=Base, y=Median, group=sample)) +
    geom_line() +
    geom_line(aes(x=Base, y=Lower_Quartile, group=sample, col="lower_quantile")) +
    geom_line(aes(x=Base, y=Upper_Quartile, group=sample, col="upper_quantile")) +
    geom_line(aes(x=Base, y=X10th_Percentile, group=sample, col="10%_quantile")) +
    facet_wrap(~sample) +
    ylim(0,45) +
    scale_color_brewer("metrics",palette = "Set1") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5))+
    scale_x_discrete(breaks=number_ticks(10))
```

## Read length distribution

This should ideally be uniform across all samples with one distinct read length dominating the distribution.

```{r read-size}
qual = read.table(file.path(path_results, "fastqc", "Sequence_Length_Distribution.tsv"), header=T, sep= "\t", check.names = F, colClasses=list("sample"="character"))
qual = qual %>% group_by(sample) %>% mutate(total=sum(Count), pct=Count/total)

ggplot(qual , aes(x=Length, y=Count, group=sample)) +
    geom_line(size=2, alpha=0.5) +
    scale_color_brewer(palette = "Set1") +
    theme_bw() +
    facet_wrap(~sample) +
    labs(y="# of reads")+
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5)) +
    scale_x_discrete(breaks=number_ticks(5))
```

The following table lists for each sample the read lengths contributing more than 10% of that sample's total number of reads:

```{r table-size, results='asis'}
kable(qual %>% filter(pct > 0.10) %>% dplyr::select(Length, sample, pct) %>% spread(Length, pct), align="c", digits=2)

```

## Read GC content

For re-sequencing projects the GC nucleotide content of sequenced reads should follow the genome distribution. Here we are checking if there are any samples where reads outside of a 10-90% GC content contribute more than 5% of the overall reads. These cutoffs can be tweaked as needed based on the G/C distribution graph:

```{r table-gc, results='asis'}
qual = read.table(file.path(path_results, "fastqc", "Per_sequence_GC_content.tsv"), header=T, sep= "\t", check.names = F, colClasses=list("sample"="character"))
qual = qual %>% group_by(sample) %>% mutate(total=sum(Count), pct=Count/total)

ggplot(qual, aes(x=GC_Content, y=pct, group=sample)) +
  geom_line(size=2, alpha=0.5) +
  scale_color_brewer(palette = "Set1") +
  theme_bw() +
  labs(y='Percent of reads', x='Percent G/C content')

qual = qual %>% filter(GC_Content>10 & GC_Content<90) %>% 
   group_by(sample) %>% summarise(pct_in_10_90=sum(pct))
kable(qual %>% filter(pct_in_10_90 <= 0.95))

```

## Read nucleotide content

Expanding on the GC read content analysis this checks for biases in nucleotide content along the position in the read. Typically, biases are introduced to due preferential (non-random) primer binding (at the beginning of the read) or other artifacts; strong biases are indicative of technical problems:

```{r read-content}
qual = read.table(file.path(path_results, "fastqc", "Per_base_sequence_content.tsv"), header=T, sep= "\t", check.names = F, colClasses=list("sample"="character"))
qual$sample = as.character(qual$sample)

qual$Base = as.numeric(qual$Base)
dd = reshape2::melt(qual, id.vars = c("Base", "sample")) %>% mutate(nt=variable)

ggplot(dd %>% filter(nt!="total"), aes(x=Base, y=value, group=sample)) +
        geom_line(size=2, alpha=0.5) +
        theme_bw() +
        ylab("% of nucleotides") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5)) +
        ylim(10,50) +
    facet_wrap(~nt)
```

## Read nuceotide biases

To help trace down possible adapter contamination, missed barcodes or other biases we list for each sample the read positions where any nucleotide is present in either less than 10% or more than 30% of the reads:

```{r table-content, results='asis'}
kable( reshape2::melt(qual, id.vars=c("Base", "sample")) %>%
           mutate(nt=variable) %>%
           filter(value > 30 | value < 10) %>% filter(Base<20) %>%
           dplyr::select(Base, sample, nt, value) %>%
           spread(Base, value),
       align="c", digits=2)
```
