
# Coverage in variants

## Variant QC

Finally, looking at global metrics for variant calls such as the number of heterozygous calls, transition/transversion ratios, and the read coverage at sites that were called as variants. 

> Unclear whether this is limited to the target / reporting regions or global. I think in general we need to clarify where metrics are a) global, b) limited to target regions, and c) limited to coverage regions.

```{r table-variants, results='asis'}
qc$ratio_het_hom = qc$Variations_heterozygous/qc$Variations_homozygous
metrics = c("sample", "Variations_total", "Variations_in_dbSNP_pct",
            "Variations_heterozygous", "Variations_homozygous", "ratio_het_hom", "Transition/Transversion")
print(kable(qc[, metrics], align="c", digits=2))
```
### Variant coverage

Another coverage plot, this time limited to positions identified as variants. Read coverage on the X-axis (limited to 100X), percentage of variants with that coverage on the Y-axis. The red line highlights the X=13 cutoff required for somewhat reliable heterogenous variant identification in germline samples.

```{r variants-coverage}
fns = list.files(file.path(path_results, "variants"), full.names = TRUE, pattern = "gc-depth-parse.tsv")
tab = data.frame()
for (fn in fns){
    dt = read.table(fn, header=T,sep="\t")
    dt = dt %>% filter(!grepl("[::.::]",depth))
    dt[,2] = as.numeric(dt[,2])
    q = quantile(dt[,2],c(0,.10,.25,.50,.75,.90,1))
    labels=factor(rev(names(q)),levels=c("0%","10%","25%","50%","75%","90%","100%"))
    dt = data.frame(variants_pct=labels, depth=q, sample=dt$sample[1])
    tab = rbind(tab, dt)
}

ggplot(tab, aes(x=depth, y=variants_pct, group=sample)) +
    geom_line(size=2, alpha=.5) +
    geom_vline(xintercept=13, color='red') +
    theme_bw() +
    xlim(0,100) +
    labs(list(x="# of reads", y="% variants with more than X reads", title="variants coverage"))
```

### Variant G/C bias

Checking for G/C bias for called variants. This is again capped at 100X coverage; X-axis is G/C content of a read, Y-axis the number of reads supporting a variant, colour reflects the number of variants at that coverage and G/C window. 


```{r variants-coverage-gc, fig.width=15, fig.height=15}
rf <- colorRampPalette(rev(brewer.pal(11,'Spectral')))
r <- c("white", rf(9))
list_p = list()
fns = list.files(file.path(path_results, "variants"), full.names = TRUE, pattern = "gc-depth-parse.tsv")
for (fn in fns[1:2]){
    dt = read.table(fn, header=T,sep="\t")

    dt = dt %>% filter(!grepl("[::.::]",depth))
    dt[,2] = as.numeric(dt[,2])
    sample = dt$sample[1]
    p = ggplot(dt, aes(CG, depth)) +
        stat_bin2d() +
        ylab("# of reads") +
        scale_fill_gradientn(colours = r, guide = FALSE) + 
        theme_bw() +
        ylim(0, 100) +
        scale_x_discrete(breaks=seq(0,100,10)) +
        ggtitle(sample)

    list_p[[as.character(sample)]]=p
}

do.call(grid.arrange, list_p)
```
